Precise spiking motifs in neurobiological and neuromorphic data

### 2022-11-09 manuscript.tex
6

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}\label{sec:discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Summary}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%: >>>>>>>> Laurent is here <<<<<<<<<<<<<<
In this review paper, we have presented recent evidence for the role of precise spiking motifs in neuroscience. In particular, we have shown that such particular motifs may play a crucial role in neurobiology, that they may be understood at the theoretical and computational levels and that they may have numerous applications in neuromorphic engineering. In particular, we have shown that:
\begin{itemize}
  \item  the efficiency of neural systems, and in particular the visual system, imposes strong constraints on the structure of neural activity which highlight the importance of precise spike times,
  \item  that growing evidence from neurobiology proves that neural systems are more than integrators and may use synchrony detection in different forms: synfne chains, traveling waves on arbitrary spiking motifs,
  \item  that many theoretical models already exist taking into account the specificity of spiking motifs, notably by using heterogeneous delays,
  \item  ultimately that using precise puking motifs could be a key ingredient in men nomorphic systems to reach similar efficiencies as biological neural systems.
\end{itemize}
 Overall, our reviewing effort has shown that a growing community is focusing on that aspect. This community is based on solid and validated evidence which is getting novel grounds thanks to current advances in technical advances. Moreover, we have also shown that this community is highly diverse, ranging from \emph{in vivo}, \emph{in computo} and \emph{in silico}. As a consequence, the global effort is still largely scattered, which limits its larger acceptance in neuroscience.

\subsection{Limits}
The different models of spike motif detection and learning that we have presented at different levels (neurobiological, theoretical, neuromorphic) individually present limitations that prevent their widespread application in neuroscience.

@@

Moreover, the learning of patterns is often done in a supervised way. Indeed, the problem of detecting polychronous groups implies to infer both the address and the precise time of occurrence of these motifs. Most of the models we have presented are based on the assumption that at least one of its variables is known: either the pattern, its identity, or its time of occurrence. This constraint is to be put in parallel with the way a biological nervous system works in which learning is done autonomously, i.e. without supervision. However, we can note that some models can perform such learning, but only in the case of data for which the motifs are easily separable.

A final limitation of the models we have presented is that they consist of a single processing layer that links an input to an output. However, we have seen that the neurobiological system uses processing loops within hierarchical graphs.

%Clustering and control for adaptation uncovers time-warped spike time patterns in cortical networks \textit{in vivo}~\citep{isbister_clustering_2021} They discovered (\textit{in vivo}) that the times that individual neurons spiked within a sequence depended on the excitability of the cortex when a stimulus was presented.
% dynamical models
%Dumas and colleagues~\citep{panahi_generative_2021} : three levels / fourth paradigm~\citep{tolle_fourth_2011} i.e., data exploration in which the scientific models are fit to the data by learning algorithms.
%Emergence of electronic architectures specialized in Sparse Event-Based Convolutions~\citep{di_mauro_alfio_sne_2022} or of sparsity aware algorithms~\citep{yin_sata_2022}.

%and oscillations? Journal Article DOI: 10.1007/BF00202899 Coherent oscillations: A mechanism of feature linking in the visual cortex?

% development
%scaffolding of neural assemblies / existence of critical periods :~\citep{dard_rapid_2022}

% todo: implement recurrent predictive rules, such as those identified to dissociate this problem~\citep{perrinet_motion-based_2012}. it may be difficult but solutions exist  ~\citep{foss_multistability_2000,perrinet_active_2014}

%hierarchies

%developpment

\subsection{Perspectives}

The limits that we have presented can be treated individually in each model as evidenced by individual efforts which try he orecome them. However, to propose a real breakthrough, we believe that future venues should provide with a unified, inter-disciplinary approach, with applications to real-word, ecological scenarios and with open and reproducible methods.

@@

Ultimately, the community should encourage the adoption of open, reproducible science. Indeed, the different models that we have displayed often come with the tools necessary he reproduce the results obtained. This is true in neurobiology~\citep{bellec_fitting_2021}, in theoretical neuroscience~\citep{izhikevich_polychronization_2006} or in engineering~\citep{grimaldi_robust_2022}. This aspect is essential to foster the emergence of inter- disciplinary projects, such as model-driven neurobiological experiments or biologically-inspired neuromorphic engineering. As such, this review aimed at paving the way to openly share the variety of resources and to offer an unified view on the role of precise spiking motifs in neuroscience.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\supplementary{The following supporting information can be downloaded at:  \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}\label{sec:discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Summary}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%: >>>>>>>> Laurent is here <<<<<<<<<<<<<<
In this review paper, we have presented recent evidence for the role of precise spiking motifs in neuroscience. In particular, we have shown that such particular motifs may play a crucial role in neurobiology, that they may be understood at the theoretical and computational levels and that they may have numerous applications in neuromorphic engineering. In particular, we have shown that:
\begin{itemize}
  \item  the efficiency of neural systems, and in particular the visual system, imposes strong constraints on the structure of neural activity which highlight the importance of precise spike times,
  \item  growing evidence from neurobiology proves that neural systems are more than integrators and may use synchrony detection in different forms: synfne chains, traveling waves on arbitrary spiking motifs,
  \item  many theoretical models already exist taking into account the specificity of spiking motifs, notably by using heterogeneous delays,
  \item  ultimately that using precise puking motifs could be a key ingredient in neuromorphic systems to reach similar efficiencies as biological neural systems.
\end{itemize}
 Overall, our reviewing effort has shown that a growing community is focusing on that aspect. This community is based on solid and validated evidence which is getting novel grounds thanks to current technical advances. Moreover, we have also shown that this community is highly diverse, operating on \emph{in vivo}, \emph{in computo} and \emph{in silico} systems. As a consequence, the global effort is still largely scattered, which limits its larger acceptance in neuroscience.

\subsection{Limits}
Additionnally, the different models of spike motif detection and learning that we have presented at these different levels (neurobiological, theoretical, neuromorphic) individually present limitations that prevent their widespread application in neuroscience.

@@

Moreover, the learning of patterns is often done in a supervised way. Indeed, the problem of detecting polychronous groups implies to infer both the address and the precise time of occurrence of these motifs. Most of the models we have presented are based on the assumption that at least one of its variables is known: either the pattern, its identity, or its time of occurrence. This constraint is to be put in parallel with the way a biological nervous system works in which learning is done autonomously, i.e. without supervision. However, we can note that some models can perform such learning, but only in the case of data for which the motifs are easily separable. More generally, to reproduce the efficiency of biological systems, one should account for the different temporal scales of adaptation, from seconds to years. For instance, the scaffolding of neural assemblies seem to follow critical periods during development~\citep{dard_rapid_2022}.

A final limitation of the models we have presented is that they consist of a single processing layer that links an input to an output. However, we have seen that the neurobiological system uses processing loops within hierarchical graphs. In general, these systems are bidirectionnally connected across different layers (for instance cortical areas), but also within a layer, as was for instance used by~\citet{izhikevich_polychronization_2006}. The whole system forms a dynamical model which may be considered globally during the learning phase, yet while taking account the constraints of the system, for instance the lack of a global clock, or the cost of fully-connected topologies.
%
\subsection{Perspectives}
%
The limits that we have presented can be treated individually in each model as evidenced by individual efforts which try he orecome them. However, to propose a real breakthrough, we believe that future venues should provide with a unified, inter-disciplinary approach, with applications to real-word, ecological scenarios and with open and reproducible methods.

@@

Ultimately, the community should encourage the adoption of open, reproducible science. Indeed, the different models that we have displayed often come with the tools necessary he reproduce the results obtained. This is true in neurobiology~\citep{bellec_fitting_2021}, in theoretical neuroscience~\citep{izhikevich_polychronization_2006} or in engineering~\citep{grimaldi_robust_2022}. This aspect is essential to foster the emergence of inter-disciplinary projects, such as model-driven neurobiological experiments or biologically-inspired neuromorphic engineering. Solutions exist to optimize these collaborations~\citep{panahi_generative_2021} and suggest the emergence of a novel paradigm for scientific advances in neuroscience~\citep{tolle_fourth_2011} i.e., by using data exploration in which the scientific models are fit to the data by learning algorithms. As such, this review aimed at paving the way to openly share the variety of resources and to offer an unified view on the role of precise spiking motifs in neuroscience.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\supplementary{The following supporting information can be downloaded at:  \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.}
